---
title: "Regression-Model 2"
output: 
  html_document: 
    code_folding: hide
    toc: true
    toc_float: true
---

```{r setup, echo = FALSE, message = FALSE}
library(tidyverse)
library(readr)
library(readxl)
library(kableExtra)
library(knitr)
library(modelr)
library(plotly)

knitr::opts_chunk$set(
  fig.width = 10,
  fig.asp = .6,
  out.width = "90%",
  message = FALSE,
  warning = FALSE
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

options(knitr.table.format = "html")
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


```{r, echo=FALSE}
main_df <- 
  read_csv("./data/maindata.csv")

main_df <- 
  main_df %>% 
  mutate(year = lubridate::year(date_of_death)) %>% 
  filter(year != 2021)
```

```{r, echo=FALSE}
pop_stat <- 
  read_csv("data/census.csv")
```

```{r, echo=FALSE}
gun_owner <- 
  read_csv("data/Gun Ownership by State 2021.csv") %>% 
  janitor::clean_names() %>% 
  select(-total_guns)
```

```{r, echo=FALSE}
crime_stat <- 
  read_excel("data/reported-violent-crime-rate-in-the-us-2020-by-state.xlsx", 
             sheet = "Data",
             range = "B5:C57") %>%
  rename(state = 1, crime_per_10e6 = 2) %>% 
  filter(state != "United States") %>% 
  mutate(crime_per_10e6 = as.numeric(crime_per_10e6))
```

```{r, echo=FALSE}
unempolyment_stat <- 
  read_excel("data/state-unemployment-rate-in-the-us-2020.xlsx", 
             sheet = "Data",
             range = "B5:C56") %>% 
  rename(state = 1, unemploy_rate = 2) %>% 
  mutate(unemploy_rate = as.numeric(unemploy_rate)/100)
```


---------------------------------------------------

### Description

Searching for more potential predictors and finding the association between innocent death and other factors, we decide to concern about the variables (Gun ownership rates, crime rate, unemployment rate in each state) and just concentrate on 2020 but in order to avoid the collinearity, we drop the highly potential correlated variable (State). 

+ In this model, predictors and variables are below:
  + **innocent death(per 10K):** count of innocent death divided by state population and multiplied by 10K
  + **age_bin:** age divided into 6 groups(0-14, 15-24, 25-34, 35-54, 55-85, 85+)
  + **gender:** Male, Female
  + **race:** usually based on visual evidence or official reports(European-American/White, African-American/Black, Native American/Alaskan, Asian/Pacific Islander, Hispanic/Latino)
  + **Gun ownership rate:** rate of the act of owning a gun, either legal or illegal
  + **Crime Rate per 100K:** be calculated by dividing the total number of reported crimes of any kind by the total population and multiplied by 100K
  + **Unemployment rate:** measures unemployment within the civilian non-institutional population aged 16 years and olde

---------------------------------------------------

### Explorative Graphs

#### Gun ownership rate by state in 2020

Gun ownership rates vary by state and Montana has the highest.

```{r}
gun_plt = 
  gun_owner %>% 
  mutate(state = fct_reorder(state, gun_ownership)) %>% 
  ggplot(aes(x = gun_ownership, y = state, fill = as.numeric(state))) + 
  geom_col(alpha = 0.6, show.legend = FALSE) +
  scale_fill_gradient(low = "#77969A", high = "#516E41") +
  labs(x = "Gun Ownership Rate",
       y = "State")

ggplotly(gun_plt) %>% layout(showlegend = F)
```

#### Crime rate by state in 2020

The crime rate varies from state to state, with Alaska showing the highest rate

```{r}
crime_plt =
  crime_stat %>%
  filter(state != "District of Columbia") %>%
  mutate(state = fct_reorder(state, crime_per_10e6)) %>%
  ggplot(aes(x = crime_per_10e6, y = state, fill = as.numeric(state))) +
  geom_col(alpha = 0.6, show.legend = FALSE) +
  scale_fill_gradient(low = "#77969A", high = "#516E41") +
  labs(x = "Crimte rate per 100k",
       y = "State")

ggplotly(crime_plt) %>% layout(showlegend = F)
```

#### Unemployment rate by state in 2020

Unemployment rates in each state show a diversified trend and Nevada has the highest unemployment rate.

```{r}
unemployment_plt =
  unempolyment_stat %>%
  filter(state != "District of Columbia") %>%
  mutate(state = fct_reorder(state, unemploy_rate)) %>%
  ggplot(aes(x = unemploy_rate, y = state, fill = as.numeric(state))) +
  geom_col(alpha = 0.6, show.legend = FALSE) +
  scale_fill_gradient(low = "#77969A", high = "#516E41") +
  labs(x = "Unemployment rate",
       y = "State")

ggplotly(unemployment_plt) %>% layout(showlegend = F)
```

---------------------------------------------------

### Statistical Analysis

```{r include=FALSE}
reg_df1 <- 
  main_df %>% 
  group_by(year, state, age_bin, gender, race) %>% 
  summarize(count = n()) %>% 
  filter(year %in% c(2010:2020), state != "DC") %>% 
  left_join(pop_stat, by = c("year", "state", "age_bin", "gender", "race")) %>% 
  mutate(innocent_kill_per100k = count / population * 10^5)

# Find distribution of the target variable `innocent_kill_per100k`
reg_df1 %>% 
  ggplot(aes(x = innocent_kill_per100k)) + 
  geom_density(color = "dodgerblue1", fill = "skyblue2") + 
  geom_vline(xintercept = mean(reg_df1$innocent_kill_per100k), linetype = "dotted")

# After transformation
reg_df1 %>% 
  ggplot(aes(x = log(innocent_kill_per100k))) + 
  geom_density(color = "dodgerblue1", fill = "skyblue2") + 
  geom_vline(xintercept = mean(log(reg_df1$innocent_kill_per100k)), linetype = "dotted")
```

```{r include=FALSE}
# Model 2 -- year 2020
# Join data
pop_gun_crime_unemply <- 
  gun_owner %>% 
  left_join(crime_stat, by = "state") %>% 
  left_join(unempolyment_stat, by = "state") %>% 
  mutate(state = state.abb[match(state, state.name)]) %>% 
  select(state, everything())

reg_df2 <- 
  reg_df1 %>% 
  filter(year == 2020) %>% 
  left_join(pop_gun_crime_unemply, by = "state") %>% 
  select(2:5, 8:11)
```

#### 1.Modeling

Because the distribution of innocent death per 10K is the same in model 1, here we still conduct log transformation and the formula is followed below:

$$ log(innocent \space death \space per \space 100k) = \beta_0 + \beta_1 age + \beta_2 gender + \beta_3 race + \beta_4Gun \space ownership \space rate$$ 
$$ + \beta_5Crime \space rate(per \space 100K) + \beta_6 Unemployment \space rate$$
```{r}
# Fit a model
fit2 <- lm(log(innocent_kill_per100k) ~ age_bin + gender + race + gun_ownership + crime_per_10e6 + unemploy_rate, 
    data = reg_df2)

summary(fit2) %>% 
  broom::tidy() %>% 
  knitr::kable() %>%
  kable_styling() %>% 
  scroll_box(height = "200px")

summary(fit2) %>% 
  broom::glance() %>% 
  knitr::kable() %>% 
  kable_styling()
```

According to the statistics in the table above, we could conclude that P-value is $2.2 \times 10^{-16}$ and smaller than significant level which means the association is significant in statistical aspect although crime rate seems have no significant association.

When concentrating on the table of R.squared, we could figure out that adjusted $R^2$ value is 0.62 which means that 62% of the variability in the outcome data can be explained by the model so that this value is generally considered a Moderate effect size and the fit is not as good as model1. If R-squared value 0.5 < r < 0.7 this value is generally considered a Moderate effect size *(Ref:Moore, D. S., Notz, W. I, & Flinger, M. A. (2013). The basic practice of statistics (6th ed.))*

#### 2.MLR dignostics

The four graphs is used for proving the homoscedasticity of residual. When looking deep into those graphs, we're able to conclude that all assumption is satisfied but it's not as good as model 1.

* Residuals vs Fitted: show that it has linear relationship due to the nearly equal distribution.
* Q-Q plot: illustrate residual is normality
* Scale-Location: The assumption of homoscedasticity has been checked.
* Residuals vs Leverage: shows that there's no too many outlier points and influential points in the plot.

```{r}
# Model diagnosis
par(mfrow = c(2, 2))
plot(fit2)
```

<br>

<font size="3"><a href = "#top" target = "_self">Back to top</a></font>